---
title: "FCM"
output: html_document
---

# Introduction
Workflow for preliminary Fuzzy Cognitive Mapping of Operation Remote Immunity

# Load libraries
```{r}
library(fcm)
library(here)
library (reshape2)
library (ggplot2)
library(dplyr)
library(doParallel)
library(doRNG)
library(ggnetwork)
library(network)
```

# Preliminary Run
## weight matrix
```{r}
prelim_WM <- read.csv(here("FCM/prelim_WM.csv"))

# convert to matrix
prelim_WM.mat <- prelim_WM[-1, c(-1,-2)]

# assign name
colnames(prelim_WM.mat) <- prelim_WM$Outcomes[-1]
rownames(prelim_WM.mat) <- c()

# assign nuermic
prelim_WM.mat <- dplyr::mutate_all(prelim_WM.mat, function(x) as.numeric(as.character(x)))
```

## activation vector
```{r}
prelim_AV <- read.csv(here("FCM/prelim_AC.csv"))
prelim_AV <- data.frame(t(prelim_AV))

colnames(prelim_AV)<- prelim_WM$Outcomes[-1]
rownames(prelim_AV) <- c()
```

## run inference
```{r}
## with default, 20 interation, Kosko inference rules, sigmoidal transformation, lambda = 1, e=0.001
output <- fcm.infer(prelim_AV, prelim_WM.mat)
output 
```

## visualization
```{r}
# assign iterations and values
iterations <- as.numeric(rownames(output$values))  
df <- data.frame(iterations, output$values)   
df2 <- melt(df, id="iterations")   

# focus on important concepts. Population size is a control for concept that should not change 
df2 %>% filter(variable %in% c("dose_delivered", "dose_allocated", "dose_uptaken", "vaccine_hesistency", "extreme_weather")) %>%
ggplot(aes(x=iterations, y=value, group=variable, colour=variable)) +
  geom_line(size = 0.7) + 
  theme_bw()

ggsave(here("FCM/figures/prelim_equlibrium.png"), height = 5)
```

## trying random starting points. Monte Carlo simulation
It seems that under default setting and preliminary set up, no chaotic behaviour nor initial-condition dependency is observed!
```{r error=FALSE, warning=FALSE, message=FALSE, results = "hide"}
fcm_mc <- function(){
  # generate random list of initial values
  AV_rand <- data.frame(t(runif(32, 0, 1)))
  colnames(AV_rand)<- prelim_WM$Outcomes[-1]
  
  # run inference
  invisible(capture.output(res <- fcm.infer(AV_rand, prelim_WM.mat)))
  return(res$values)
}

# parallelized, reproducible Monte Carlo simulation
cl <- makeCluster(8)
registerDoRNG(137)
registerDoParallel(cl)
fcm_mc_res <- foreach(i = 1:10000, .packages = "fcm") %dorng%{
  fcm_mc()
}

# combine all dataframe
fcm_mc.df <- bind_rows(fcm_mc_res, .id =  "iteration")

# get equilibrium of each iteration. 
fcm_mc.eq <- fcm_mc.df %>% 
  group_by(iteration) %>% 
  slice(n())

fcm_mc.eq2 <- melt(fcm_mc.eq)

# at 5 decimal place rounding, no difference between iterations!
fcm_mc_eq.ls <- fcm_mc.eq2 %>% 
  group_by(variable) %>% 
  summarize(eq = unique(round(value, 5)))
  
```

# Plot fuzzy cognitive map
```{r}
# make adjacency matrix
prelim_WM.mat2 <- prelim_WM.mat
rownames(prelim_WM.mat2) <- names(prelim_WM.mat)

# make edge list
fcm.edgels <- melt(as.matrix(prelim_WM.mat2))


# add on attributes
fcm.edgels <- cbind(fcm.edgels, type = c(rep("Outcome", 4), rep("Product", 5), rep("Allocation", 2), rep("Delivery", 6), rep("Uptake", 5), rep("Human_resource", 3), rep("Leadership", 4), rep("Demographic", 3)))

# get rid of 0 weights 
fcm.edgels <- fcm.edgels %>% filter(value != 0)

# get network object
fcm.net <- graph.data.frame(fcm.edgels, directed = T)

# get list of color for vertex
color.ls <- scales::hue_pal()(8)

color.df <- left_join(data.frame(type = E(fcm.net)$type), 
          data.frame(type = unique(E(fcm.net)$type), color = color.ls))

# compute degree
deg <- degree(fcm.net, mode="all")

# get list of color for edge
edge_color <- fcm.edgels %>% 
  mutate(edge_color = ifelse(value < 0, "red", "grey"))


# get better layout
layout <- layout.fruchterman.reingold(fcm.net, niter=5000, area = vcount(fcm.net)^4*10)

# plot
tiff(here("FCM/figures/fcm_prelim.tiff"), width = 1000, height = 1000, res = 120)
plot(fcm.net, 
     edge.width = abs(E(fcm.net)$value)*3, 
     layout = layout,
     edge.arrow.size = 0.4,
     vertex.size = deg,
     vertex.color = color.df$color,
     vertex.label.font = 1,
     vertex.label.color = "black",
     vertex.frame.color = "transparent",
     edge.color = edge_color$edge_color
     )
dev.off()
```



