---
title: "WAHA-SLFNHA_demography"
output: html_document
runtime: shiny
self_contained: true
---
# Introduction
Code for visualization of First Nations communities served by the Weeneebayko Area Health Authority (WAHA) and the Sioux Lookout First Nations Health Authority (SLFNHA). All data is compiled from the 2016 Canadian Census.

## Data Processing
### Load packages
```{r, message=FALSE, results='hide'}
library(dplyr)
library(tmap)
library(sf)
library(cancensus)
library(here)
library(stringr)
library(rgdal)
library(ggplot2)
library(shiny)
library(tmaptools)

# load API key for Cancensus. Replace with custom
api_key <- readr::read_file(here("comparative", "census_api.txt"))
options(cancensus.api_key = api_key)
```

### Process Data
#### Retrieving SGC of communities served
```{r}
# list of communities served by WAHA
waha_com <- c("Attawapiskat", "Fort Albany", "Kashechewan", "Moosonee", "Weenusk", "Moose Cree")

# list of communities served by SLFNHA
slf_com <- read.delim(here("comparative","slf_com.txt"), header = F)$V1
```

##### Using alternative sgc retrieved from https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/abpopprof/about-apropos/tab-band-bande.cfm?LANG=E

WAHA
```{r}
# read in the file
indigenous_sgc <- read.csv(here("comparative","first_nation_sgc.csv"))

# reassign name
names(indigenous_sgc) <- c("Province", "Location", "census_sgc")

# Moosonee not found
waha_sgc <- indigenous_sgc %>% 
  filter(str_detect(Location, paste(waha_com, collapse = "|"))) %>%
  filter(Province == "Ontario") %>% 
  select("census_sgc")

# get list of sgcs
waha_sgc2 <- unlist(lapply(waha_sgc$census_sgc, function(x){
  y <- str_split(x, ";")
  lapply(y, function(z) sub(".*\\((\\d+)\\).*", "\\1", z))
}))

# manually add Moosonee sgc
waha_sgc3 <- append(waha_sgc2, 3556106)
```

SLFNHA
```{r}
# Koocheching, McDowell Lake not found, 
slf_sgc <- indigenous_sgc %>% 
  filter(str_detect(Location, paste(slf_com, collapse = "|"))) %>%
  filter(Province == "Ontario") %>% 
  select("census_sgc")

# get list of sgcs
slf_sgc2 <- unlist(lapply(slf_sgc$census_sgc, function(x){
  y <- str_split(x, ";")
  lapply(y, function(z) sub(".*\\((\\d+)\\).*", "\\1", z))
}))

# manually add in:
## Pikangikum is part of independent First Nations: 3560077
# Wawakapewin (Long Dog Lake): 3560100
# Weagamow Lake 87: 3560059
slf_sgc3 <- append(slf_sgc2, c(3560077, 3560100, 3560059))
```


### Retrieving Census data
```{r, message=FALSE,results='hide'}
waha_census <- get_census(dataset = "CA16",
           regions = list(CSD = waha_sgc3),
           level = "CSD",
           vectors = c("v_CA16_401", "v_CA16_404", "v_CA16_406", "v_CA16_407", "v_CA16_4", "v_CA16_61","v_CA16_244", "v_CA16_379", "v_CA16_425", "v_CA16_4872", "v_CA16_4870","v_CA16_4856", "v_CA16_4858", "v_CA16_512"),
           geo_format = "sf")

slf_census <- get_census(dataset = "CA16",
           regions = list(CSD = slf_sgc3),
           level = "CSD",
           vectors = c("v_CA16_401", "v_CA16_404", "v_CA16_406", "v_CA16_407", "v_CA16_4", "v_CA16_61","v_CA16_244", "v_CA16_379", "v_CA16_425", "v_CA16_4870", "v_CA16_4872","v_CA16_4856", "v_CA16_4858", "v_CA16_512"),
           geo_format = "sf")
```

# Prepare for plotting
```{r,results='hide'}
# Assign health district
slf_census$health <- "SLFNHA"
waha_census$health <- "WAHA"
waha_slf.census <- rbind(slf_census, waha_census)

# make numeric
waha_slf.census <- waha_slf.census %>% 
  mutate_at(c(8,12:25), as.numeric)

# calculate % statistics for overcrowding and major repair. Definition of overcrowding from Stat Canada 2016 Aboriginal Profile
waha_slf.census2 <- waha_slf.census %>% 
  dplyr::mutate("Major repairs needed (%)" = `v_CA16_4872: Major repairs needed`/`v_CA16_4870: Total - Occupied private dwellings by dwelling condition - 25% sample data`*100,
         "Overcrowded (%)" = `v_CA16_4858: More than 1 person per room`/`v_CA16_4856: Total - Private households by number of persons per room - 25% sample data`*100)
```

# Append indigenous governance structure
```{r,results='hide'}
# load in info retrieved form http://firstnation.ca/communities
tribal <- read.csv(here("comparative", "tribal_council.csv"))

# fuzzy match what we can
ci_str_detect <- function(x, y){str_detect(x, regex(y, ignore_case = TRUE))}

waha_slf.census3 <- fuzzy_left_join(waha_slf.census2, tribal, match_fun = ci_str_detect, by = c("name" = "community"))

# manual data entry
waha_slf.census3 %>% select(name, community)

waha_slf.census4[2, c(31:36)] <- c(258, "Windigo First Nations Council", "Grand Council of Treaty 3",	"Treaty 3",	50.420986,	-90.658836)
waha_slf.census4[3, c(31:36)] <- c(203,	"Independent First Nations Alliance",	"Nishnawbe-Aski Nation",	"Treaty 9",	51.232258,	-90.234833)
waha_slf.census4[5, c(31:36)] <- c(259,	"Windigo First Nations Council",	"Nishnawbe-Aski Nation",	"Treaty 9",	51.1492,	-91.613617)
waha_slf.census4[6, c(31:36)] <- c(186,	"Matawa First Nations",	"Nishnawbe-Aski Nation",	"Treaty 9",	51.634003,	-85.920296)
waha_slf.census4[7, c(31:36)] <- c(183,	"Matawa First Nations",	"Nishnawbe-Aski Nation",	"Treaty 9",	51.558786,	-87.905557)
waha_slf.census4[9, c(31:36)] <- c(203,	"Independent First Nations Alliance",	"Nishnawbe-Aski Nation", "Treaty 9",	51.232258,	-90.234833)
waha_slf.census4[12, c(31:36)] <- c(204,	"Windigo First Nations Council",	"Nishnawbe-Aski Nation",	"Treaty 9",	52.949673,	-91.344109)
waha_slf.census4[14, c(31:36)] <- c(237,	"Keewaytinook Okimakanak",	"Nishnawbe-Aski Nation",	"Treaty 9",	52.625144,	-94.049149)
waha_slf.census4[16, c(31:36)] <- c(209,	"Independent First Nations Alliance",	"Independent",	"Treaty 9",	53.82214,	-89.894257)
waha_slf.census4[22, c(31:36)] <- c(239,	"Matawa First Nations",	"Nishnawbe-Aski Nation",	"Treaty 9",	52.202281,	-88.020213)
waha_slf.census4[24, c(31:36)] <- c(217,	"Shibogama First Nations Council",	"Nishnawbe-Aski Nation",	"Treaty 9",	52.931259,	-89.299622)
waha_slf.census4[25, c(31:36)] <- c(241,	"Matawa First Nations",	"Nishnawbe-Aski Nation",	"Treaty 9",	52.743681,	-88.51272)
waha_slf.census4[26, c(31:36)] <- c(206,	"Shibogama First Nations Council",	"Nishnawbe-Aski Nation", "Treaty 9",	53.831459,	-89.568787)
waha_slf.census4[27, c(31:36)] <- c(239,	"Matawa First Nations",	"Nishnawbe-Aski Nation",	"Treaty 9",	52.202281,	-88.020213)
waha_slf.census4[30, c(31:36)] <- c(239,	"Matawa First Nations",	"Nishnawbe-Aski Nation",	"Treaty 9",	52.202281,	-88.020213)
waha_slf.census4[32, c(31:36)] <- c(234,	"Shibogama First Nations Council",	"Nishnawbe-Aski Nation",	"Treaty 9",	53.450237,	-89.143066)
waha_slf.census4[c(35), c(31:36)] <- c(144,	"Mushkegowuk Council",	"Nishnawbe-Aski Nation",	"Treaty 9",	51.261269,	-80.59536)
waha_slf.census4[c(36), c(31:36)] <- c(144,	"Mushkegowuk Council",	"Nishnawbe-Aski Nation",	"Treaty 9",	51.261269,	-80.59536)
waha_slf.census4[40, c(31:36)] <- c(146,	"Mushkegowuk Council",	"Nishnawbe-Aski Nation",	"Treaty 9",	55.0108,	-85.423164)

write_sf(waha_slf.census4, here("comparative", "waha_slf_census4.shp"))
```

# Generate bounding box for the 2 health authorities (Tribal Council)
```{r}
waha_bb <- bb_poly(waha_slf.census2 %>% filter(health == "WAHA"))

slf_bb <- bb_poly(waha_slf.census2 %>% filter(health == "SLFNHA"))
```

# visualize Governance structure (Primary Organization)
```{r}
tm_shape(st_as_sf(waha_slf.census4)) +
  tm_dots(col = "primary_org",
          id = c("name"),
          size = 0.07) +
  tm_shape(waha_bb) + tm_borders() +
  tm_shape(slf_bb) + tm_borders()
```

# Visualize governance structure (Tribal council)
```{r}
tm_shape(st_as_sf(waha_slf.census4)) +
  tm_dots(col = "tribal_council",
          id = c("name"),
          size = 0.07) +
  tm_shape(waha_bb) + tm_borders() +
  tm_shape(slf_bb) + tm_borders()
```

# visualize demographic differences
```{r}
world_vars <- names(st_drop_geometry(waha_slf.census4[c(8,12:21, 29,30)]))

ui <- fluidPage(
  tmapOutput("map"),
  selectInput("var", "Variable", world_vars)
)

server <- function(input, output, session) {
  output$map <- renderTmap({
  tm_shape(waha_slf.census2) +
  tm_dots(col = input$`var`, 
          palette = "viridis",
          style = "cont",
          size = 0.125,
          alpha = 0.7,
          id = c("name"),
          colorNA = "grey",
          textNA = "Missing",
popup.vars = c("Health Authority:" = "health",
               "Variable:" = input$`var`), title = gsub(".*: ", "", input$`var`)) +
      tm_shape(waha_bb) + tm_borders() +
      tm_shape(slf_bb) + tm_borders() +
      tm_scale_bar() +
      tm_compass(type="radar", position=c("left", "top"), show.labels = 3)

  })
}	

shinyApp(ui, server,  options = list(height = 800))
```

## Visualizing demography differences
```{r}
# filter out region with no population data
waha_slf.census3 <- waha_slf.census2 %>% filter(Population != 0) %>%
  st_drop_geometry()

# rename variables 
names(waha_slf.census3) <- gsub(".*: ", "", names(waha_slf.census3))

# gather
waha_slf.census4 <- waha_slf.census3 %>% 
  tidyr::gather(key, value, -health) %>% 
  filter(key %in% names(waha_slf.census3)[12:29])

waha_slf.census3$`Population, 2016`
# filter variables that are more interesting
var_int <- c("Population, 2016", "Population density per square kilometre", "Major repairs needed (%)", "Overcrowded (%)", "0 to 14 years", "65 years and over")

# Violine plot with overlaid points
waha_slf.census4 %>% 
  filter(key %in% var_int) %>% 
  ggplot(aes(x = health, y = as.numeric(value), fill = as.factor(health))) +
  geom_violin(trim = F) +
  stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="black") +
  facet_wrap(~key, scales = "free", labeller = label_wrap_gen(width = 25)) +
  labs(fill = "Health authority", y = "Value", x = "") +
  geom_point(position=position_jitterdodge(), alpha=0.3) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

ggsave(here("figures/waha_slf_demo_violin.png"), width = 7, height = 6)
```

